---
title: "Session 2: Homework 1"
author: "Group 1"
date: "`r Sys.Date()`"
output:
  html_document:
    theme: flatly
    highlight: zenburn
    number_sections: yes
    toc: yes
    toc_float: yes
    code_folding: show
---


```{r, setup, echo=FALSE}
knitr::opts_chunk$set(
  message = FALSE, 
  warning = FALSE, 
  tidy=FALSE,     # display code as typed
  size="small")   # slightly smaller font for code
options(digits = 3)

# default figure size
knitr::opts_chunk$set(
  fig.width=6.75, 
  fig.height=6.75,
  fig.align = "center")
```


```{r load-libraries, warning=FALSE, message=FALSE, echo=FALSE}
library(tidyverse)
library(mosaic)
library(ggthemes)
library(lubridate)
library(fivethirtyeight)
library(here)
library(skimr)
library(janitor)
library(vroom)
library(tidyquant)
library(lubridate)

party_colors <- c("#2E74C0", "#CB454A") #color codes for the ggplot

```


# Challenge 2: 2016 California Contributors

In this post we are going to examine contributions made to Hillary Clinton and Donald Trump in the state of California before the 2016 US primary elections. The traditionally blue state went was clearly won by the democrats and their nominee Hillary Clinton. 




We start with loading the data.

```{r, load_CA_data, warnings= FALSE, message=FALSE, cache=FALSE}

CA_contributors_2016 <- vroom::vroom(here::here("data","CA_contributors_2016.csv")) # Loading Contribution Data Set

zip_codes_data <- vroom::vroom("http://www.uszipcodelist.com/zip_code_database.csv")

# inspecting the data sets
glimpse(CA_contributors_2016) 

# inspecting the data sets
glimpse(zip_codes_data)

#getting rid of unessecary data, only keeping the zip code and corresponding city
zip_codes <- zip_codes_data %>% 
  select(zip, primary_city)

```
Merging the two data sets:

While glimpsing at the two data sets, we noticed that the while both data set have the column, "zip", one is saved as a double and one as a character.

In order to merge the two with left_join, we have to change one of the two. Here we change the zip variable in the CA_contributors_2016 data set from double to character
```{r, merging_data}

CA_contributors_2016_amd <- CA_contributors_2016 %>%
  mutate(zip = as.character(zip))

# We can now merge the two datasets

new_dataset <- left_join(CA_contributors_2016_amd, zip_codes, by = "zip") #joining the two data set using the variable "zip" which is present in both datasets
glimpse(new_dataset)

```



## Cumulative earnings

```{r}

cummulative_contr <-  new_dataset %>% 
  group_by(cand_nm, contb_date) %>% 
  summarize(daily_contr_sum = sum(contb_receipt_amt)) %>% 
  mutate(daily_contr_cumsum = cumsum(daily_contr_sum))
  NULL
 
View(cummulative_contr)  

glimpse(cummulative_contr) 


test <- cummulative_contr %>% 
  filter(cand_nm == "Clinton, Hillary Rodham" | cand_nm == "Trump, Donald J.",
         contb_date <= as.Date("2016-11-08"))

test
  


```

```{r}

ggplot(test, aes(contb_date, daily_contr_cumsum/1000000, color = cand_nm)) +
  geom_line(size = 1) + 
  theme_bw() +
  labs(title = "Trump starts raising funds later than Clinton in CA", subtitle = "Cummulative contributions recieved up to election day in CA", y="Cummulative contributions (in mio $)", x = NULL, color = "Candidate") +
  scale_x_date(date_labels = "%b %y") +
  scale_color_manual(values = party_colors) +
  NULL 

```

We can see that Clinton 




As we can see we have know successfully added the city to each contribution.

Filtering for the relevant candidates:

As the graph above shows the Top 10 cities for only two candidates, we are going to filter out the rest.
  
```{r, filtering}

new_dataset_1 <- new_dataset %>% 
  filter(cand_nm == "Clinton, Hillary Rodham" | cand_nm == "Trump, Donald J.") #filter for Trump and Clinton


```

```{r, reordering}

library(tidytext)

top_10_cities_orderd <- new_dataset_1 %>% 
  group_by(primary_city, cand_nm) %>% #grouping data by candidate and city
  summarise(contribution = sum(contb_receipt_amt)) %>%  # adding up the contributions for each candidate and city
  ungroup() %>% 
  group_by(cand_nm) %>% #grouping data by candidate only
  top_n(10, contribution) %>% #selecting the top contributing cities for each group
  ungroup() %>% 
  mutate(primary_city = reorder_within(primary_city, contribution, cand_nm)) #reordering the cities by contribution for each candidate


```

Plotting the data:

```{r, plotting}



final_plot <- ggplot(top_10_cities_orderd, 
                     aes(x = contribution, #show contribution on x axis
                         y = reorder(primary_city, contribution), #shows cities on y axis, orderd by contribution
                         fill = cand_nm)) + 
  geom_col(show.legend = FALSE) + #hides the legend
  scale_x_continuous(labels = scales::dollar) + #show y-axis values in dollar
  theme_bw() + #changes the grid theme
  facet_wrap(~cand_nm, scales = "free") + #shows data in two facets, one for each candidate
  scale_fill_manual(values = party_colors) + #colors the two facets in blue and red
  labs(x = "Amount raised", y = NULL, subtitle = "Where did candidates raise the most money?") + #formats lables and plot titles
  theme(plot.subtitle = element_text(face = "bold")) + #makes plot title bold
  theme(axis.title.x = element_text(face = "bold")) + #makes x-axis title bold
  scale_y_reordered() + #necessary to properly show y-axis lables when "reorder_within" function is used
  theme(aspect.ratio = 1) + #changes facets to appear as squares
  NULL

final_plot
```






## Average contribution per state





